### 2012-08-17 plan/plan.tex
Started on the plan!

\documentclass{book}
\usepackage{amsmath}
% \usepackage{hyperref}

\input ../defs.tex
\newcommand{\T}{^\dagger}
\newcommand{\reals}{{\mbox{\bf R}}}
\newcommand{\comps}{{\mbox{\bf C}}}

\title{Objective-First Nanophotonic Design Plan} 
\author{Jesse Lu}
\begin{document}
\maketitle
\tableofcontents

\chapter{Introduction}
\section{Problem statement}
Our goal is to create a software package 
    to enable the design of nanophotonic devices.
Presently, almost all nanophotonic devices are designed
    by guessing what a good design might look like,
    based on intuition and experience,
    and then optimizing it by trial-and-error.
In contrast, we want to enable the \emph{inverse design}
    of nanophotonic devices;
    that is, to design devices by simply describing the 
    desired performance that it must achieve.

To put things mathematically, we want to create software 
    to solve the following problem,
    \EE {the plan (single mode)}
        {\minimize&  f(x) + g(z) \\
        \subto&     A(z)x - b(z) = 0}
    where 
    \BI $x$ and $z$ are the variables representing 
            the field device our device produces and
            the structure of the device respectively,
    \I  $f(x)$ and $g(z)$ are our \emph{design objectives},
            which tell us the desirable properties 
            that we would like our device to achieve, and
    \I  $A(z)x - b(z)$ is the \emph{physics residual},
            the underlying physical laws which must be met. \EI

In general, we need to consider multiple fields produced 
    by the device.
Our problem statement is then
    \EE {the plan (multi-mode)}
        {\minimize&  \sum_i^N f_i(x_i) + g(z) \\
        \subto&     A_i(z)x_i - b_i(z) = 0,\quad\text{for $i = 1, \ldots, N.$}}

\section{Capabilities of our software}
We have implemented various 
    \emph{optimization paradigms} and \emph{structure parameterizations}
    which a user can arbitrarily combine in order to 
    design nanophotonic devices.

Specifically, the user can choose to work in either
    a \emph{local} or \emph{global} optimization paradigm:
    \BI the local paradigm uses the adjoint method
            to find small changes in the structure which
            will decrease the design objective;
    \I  the global paradigm uses the objective-first method
            to arrive at a structure by forcing
            the design objective to be met from the start. \EI

And the user can also choose between the various structure parameterizations 
    which include density, boundary, 
    continuous, and combinatoric parameterizations.
These parameterizations describe the structure as either
    \BI a material density at every point in space;
    \I  a boundary between objects of different materials;
    \I  a set of continuously variable user-defined parameters; or 
    \I  a combination of user-defined objects. \EI

\chapter{Theory}
\section{Mathematically rigorous statement of the problem}
As previously stated, the problem we want to solve is 
    \EE {rigorous problem statement}
        {\minimize&  \sum_i^N f_i(x_i) + g(z) \\
        \subto&     A_i(z)x_i - b_i(z) = 0,\quad\text{for $i = 1, \ldots, N.$}}
To now be more precise,
    \BI $x_i \in \comps^m$ are the field variables,
    \I  $z \in \reals^n$ is the structure variable,
    \I  $f_i(x_i) \in \comps^m \to \reals$ are the field design objectives,
    \I  $g(z) \in \reals^n \to \reals$ is the structure design objective,
    \I  $A_i(z)x_i - b_i(z)$ are the physics residuals, with
    \I  $A_i(z) \in \comps^{m \times m}$ and
    \I  $b_i(z) \in \comps^m$. \EI

\subsection{Definition of physics residual}
The physics residual corresponds to the electromagnetic wave equation
    \E  {wave equation in E}
        {(\curl \mu^{-1} \curl - \omega^2 \epsilon) E = -i \omega J}
    which is described as $A_i(z)x_i - b_i(z)$ via
    \BI $\curl \mu^{-1} \curl - \omega^2 \epsilon \to A_i(z)$,
    \I  $\epsilon \in z$,
    \I  $E \to x_i$, and
    \I  $-i \omega J \to b_i(z)$. \EI

\subsection{Bi-affine property of the physics residual}
Critically, \eq{wave equation in E} is not only linear in $E$,
    but is also affine in $\epsilon$.
This allows us to form the following relationship,
    \E  {bi-affine}
        {A_i(z)x_i - b_i(z) = B(x_i)z - d(x_i) = 0,}
    which will prove to be very useful.

\subsection{Definition of the field design objective}
Although $f_i(x_i)$ can take on virtually any form,
    we choose to define it very specifically as
    \E  {field design objective definition}
        {f_i(x_i) = \sum_j I_+(|c_{ij}\T x_i| - \alpha_{ij})
            + I_+(\beta_{ij} - |c_{ij}\T x_i|),}
    where $c_{ij} \in \comps^m$ and 
    $I_+$ is the indicator function on nonnegative reals,
    \E{}{I_+(u) = \begin{cases} 0 & u \ge 0, \\ 
                                \infty & u < 0. \end{cases}}

Such a design objective implements the constraints 
    $\alpha_{ij} \le |c_{ij}\T x_i| \le \beta_{ij}$
    which can be interpreted physically as 
    constraining the power emitted into the optical modes
    represented by $c_{ij}$.

\subsection{Choice of the structure design objective}
In contrast to the narrow definition of the field design objective,
    the structure design objective is relatively unconstrained;
    taking on various forms to best suit the needs of
    the structure parameterization in use.

\section{Properties of the problem}
\subsection{nonconvexity}
\subsection{optimality condition}

\section{General strategy to solve the problem}
\subsection{Field update}
Give the \emph{specific} interfaces here. 
What exactly gets input and what does each update output?
\subsection{Structure update}

\end{document}