### 2013-03-28 thesis/intro.tex
Getting there with thesis.

\chapter{Introduction}
And now for an introduction...
\myfig{integrated_circuit}{}
\myfig{des_complexity}{}