### 2013-03-29 thesis/appendix2.tex

\chapter{Metamaterial-like objective-first results}\label{ob-1 meta}