### 2013-03-29 thesis/chapter3.tex

% \documentclass[letterpaper,10pt]{article}
% \usepackage{graphicx}
% \usepackage{opex3}
% \usepackage{amsmath,amssymb}
% 
% 
% \begin{document}
% \title{Objective-first design of 
%     high-efficiency, small-footprint couplers between 
%     arbitrary nanophotonic waveguide modes}
% \author{Jesse Lu$^\ast$ and Jelena Vu\v{c}kovi\'{c}}
% \address{Stanford University, Stanford, California, USA.}
% \email{jesselu@stanford.edu}
% 
% \begin{abstract}
% We present an algorithm for designing
%     high efficiency ($\sim$98\%), 
%     small-footprint (1.5-4 square vacuum wavelengths)
%     couplers between arbitrary nanophotonic waveguide modes
%     in two dimensions.
% Our ``objective-first'' method is
%     computationally fast (15 minutes on a single-core personal computer), 
%     requires no trial-and-error, and
%     does not require guessing a good starting design.
% We demonstrate designs for various coupling problems which suggest 
%     that our method allows for the design of any 
%     single-mode, linear optical device.
% \end{abstract}
% \ocis{230.75370, 130.3990.}
% 
% \begin{thebibliography}{99}
% \bibitem{fibergrating} Y. Tang, Z. Wang, L. Wosinski, U. Westergren, and S. He,
%     ``Highly efficient nonuniform grating coupler for silicon-on-insulator 
%     nanophotonic circuits,''
%     Opt. Lett. \textbf{35}, 1290-1292 (2010).
% \bibitem{ridge}  K. K. Lee, D. R. Lim, L.C. Kimerling, J. Shin, and F. Cerrina, 
%     ``Fabrication of ultralow-loss Si/SiO2 waveguides by roughness reduction,''
%     Opt. Lett. \textbf{26}, 1888-1890 (2001).
% \bibitem{pcslow} Y. A. Vlasov, M. O'Boyle, H. F. Hamann, and S. J. McNab,
%     ``Active control of slow light on a chip with photonic crystal waveguides,''
%     Nature \textbf{438}, 65-69 (2005).
% \bibitem{slotfocus} M. Lipson, 
%     ``Guiding, modulating, and emitting light on 
%     silicon-challenges and opportunities,'' 
%     J. Lightwave Technol. \textbf{23}, 4222-4238 (2005).
% \bibitem{active} J. Van Campenhout, P. Rojo Romeo, P. Regreny, C. Seassal, 
%     D. Van Thourhout, S. Verstuyft, L. Di Cioccio, J.-M. Fedeli, 
%     C. Lagahe, and R. Baets, 
%     ``Electrically pumped InP-based microdisk lasers integrated with a 
%     nanophotonic silicon-on-insulator waveguide circuit,'' 
%     Opt. Express \textbf{15}, 6744-6749 (2007).
% \bibitem{metallic} L. Tang, S. E. Kocabas, S. Latif, A. K. Okyay, 
%     D. S. Ly-Gagnon, K. C. Saraswat, and D. A. B. Miller, 
%     ``Nanometre-scale germanium photodetector enhanced by a 
%     near-infrared dipole antenna,'' 
%     Nature Photonics \textbf{2}, 226-229 (2).
% % \bibitem{fwadia} V. R. Almeida, R. R. Panepucci, and M. Lipson, 
% %     ``Nanotaper for compact mode conversion,'' 
% %     Opt. Lett. \textbf{28}, 1302-1304 (2003) 
% % \bibitem{wwadia} S. G. Johnson, P. Bienstman,  M. A. Skorobogatiy, 
% %     M. Ibanescu1, E. Lidorikis, and J. D. Joannopoulos,
% %     ``Adiabatic theorem and continuous coupled-mode theory for 
% %     efficient taper transitions in photonic crystals,''
% %     Phys. Rev. E \textbf{66}, 066608 (2002)
% % \bibitem{deriv} F. Wang, J. S. Jensen, O. Sigmund, 
% %     ``Robust topology optimization of photonic crystal waveguides with 
% %     tailored dispersion properties.'' 
% %     J. Opt. Soc. Am. B \textbf{28}, 387-397 (2011)
% % \bibitem{boydbook} S. Boyd, and L. Vandenberghe, 
% %     \emph{Convex Optimization} 
% %     (Cambridge University Press, 2004)
% \bibitem{prevwork} J. Lu, S. Boyd, and J. Vuckovic, 
%     ``Inverse design of a three-dimensional nanophotonic resonator,''
%     Opt. Express \textbf{19}, 10563-10750 (2011). 
% \bibitem{veronis} G. Veronis, and S. Fan, 
%     ``Theoretical investigations of compact couplers between dielectric slab 
%     waveguides and two-dimensional metal-dielectric-metal plasmonic waveguides,'' 
%     Opt. Express \textbf{15}, 1211-1221 (2007).
% \bibitem{yang} R. Yang, R. A. Wahsheh, Z. Lu, and M. A. G. Abushagur, 
%     ``Efficient light coupling between dielectric slot waveguide and plasmonic
%     slot waveguide,'' Opt. Lett. \textbf{35}, 649-651 (2010).
% \bibitem{code} \url{www.github.com/JesseLu/objective-first}
% \end{thebibliography}
% 
\chapter{Objective-first design} \label{ob-1}
% 
% % Why is waveguide mode conversion important?
% Optical mode conversion, 
%     the efficient transfer of photons from one guided mode to another,
%     is a fundamental requirement in nanophotonics.
% For instance, efficient conversion between waveguides modes
%     is essential for:
%     \begin{itemize}
%         \item Coupling to and from optical fiber\cite{fibergrating}, 
%         to communicate with the outside world.
%     \item Coupling between various nanophotonic waveguides, 
%         since different waveguides are best suited for different applications.
%         For example, ridge waveguides seem ideal for 
%             low-loss transport\cite{ridge},
%             but other waveguides, 
%             such as photonic crystal waveguides or slot waveguides,
%             may be better suited for slow-light\cite{pcslow} 
%             or nonlinear optical devices based on 
%             localized field intensities\cite{slotfocus}.
%     \item Coupling between different materials systems such as
%        passive, active\cite{active}, and metallic\cite{metallic} devices. 
%     \end{itemize}
% In fact, the problem of converting between nanophotonic modes is essentially
%     the function of all linear nanophotonic devices, since
%     any linear device is completely characterized by its input and output modes
%     and the coupling coefficients between them.
% 
% In this paper, we present a method to solve the problem 
%     of designing single-mode, linear nanophotonic devices.
% We then demonstrate this method by designing coupling structures 
%     between various nanophotonic waveguides.
% Furthermore, we show that our method 
%      does not require a good initial design,
%      is computationally fast,
%      and can generate high efficiency couplers within a very small footprint. 
% 
% 
% % Why are current solutions insufficient?
% % Brute force
% Of the design strategies currently available,
%     brute-force parameter search is the most often employed 
%     because of its sheer simplicity.
% Although it may be suitable for tuning existing designs,
%     the parameter space for most practical devices is simply too large
%     for such a strategy to be tractable.
% 
% % Adiabatic mode conversion (large devices, symmetry breaking)
% Adiabatic mode conversion strategies have been succesful
%     for certain fiber-waveguide\cite{fwadia} and 
%     waveguide-waveguide\cite{wwadia} couplers,
%     even if the resulting devices often require large footprints.
% However, adiabatic approaches cannot be used in many important cases 
%     such as coupling in the out-of-plane direction, or
%     coupling between modes of opposite symmetry.

% How does your algorithm solve this problem?
% \subsection{Advantages of objective-first design}
% \section{Objective-first approach}
% % What is objective-first optimization?
% Physical structures are typically designed by solving the following problem:
%     \begin{subequations}\label{eq:adj}
%     \begin{align} 
%     \text{decrease} & \quad f(x) \label{eq:adj:obj} \\ 
%     \text{subject to} & \quad g(x,p) = 0, \label{eq:adj:con}
%     \end{align}
%     \end{subequations}
%     where $x$ is the field variable and $p$ is the structure variable.
% Here, $f(x)$, the \emph{design objective}, 
%     calculates the performance of the device 
%     (e.g. amount of power lost); 
%     while $g(x,p)$ is the underlying physical equation for the system
%     (e.g. the electromagnetic wave equation).
% 
% In contrast, the objective-first approach solves 
%     \begin{subequations}\label{eq:ob1}
%     \begin{align} 
%     \text{decrease} & \quad \|g(x,p)\|^2 \label{eq:ob1:obj} \\ 
%     \text{subject to} & \quad f(x) = 0, \label{eq:ob1:con}
%     \end{align}
%     \end{subequations}
%     where $\|g(x,p)\|^2$ is the \emph{physics residual}.
% We term this formulation ``objective-first''
%     because the design objective is prioritized even above satisfying physics;
%     specifically, we force our design to always exhibit the desired performance
%     ($f(x) = 0$), even at the expense of
%     not perfectly satisfying the underlying physics which governs its operation.
% 
% The motivation behind this formulation is two-fold.
% First, this approach allows one to arrive at a locally-optimal design rapidly 
%     by allowing $x$ and $p$ to vary independently, 
%     as opposed to Eq.~\ref{eq:adj}
%     where the value of $x$ is completely dependent on $p$.
% Second, it enables an increase in the likelihood of 
%     arriving at a high efficiency design
%     by forcibly imposing $f(x) = 0$,
%     and thereby circumvents any local optima consisting of
%     low-performance devices.
% 
% How is it applied to waveguide coupler design?
\section{Problem formulation}
Our next realization was that,
    for linear nanophotonic devices,
    only certain portions of the electromagnetic field were relevant 
    to device performance--namely the boundary fields
    at the input and output ports of the device.
This realization motivated the following problem formulation:
    \begin{subequations}
    \begin{align} 
    \minimize & \|A(z)x - b\|^2 \\
    \subto & x_\text{boundary} - \hat{x}_\text{boundary} = 0 \\
        & z_\text{min} \le z \le z_\text{max},
    \end{align} \label{ob1 problem}
    \end{subequations}
    which replaces the field regularization term from \eq{iterative problem}
    with the hard constraint $x_\text{boundary} - \hat{x}_\text{boundary} = 0$.
This constraint forces the field at the device boundary ($x_\text{boundary}$)
    to be equal to that of an ideal device ($\hat{x}_\text{boundary}$),
    as shown in fig.~\ref{p3/intro}.
We call this type of formulation ``objective-first''
    since the performance of the device is prioritized above
    the physicality of the device.

\begin{figure}[htb]
    \centering
    \includegraphics[width=\textwidth]{p3/intro} 
    \caption{Boundary-value formulation of the design objective.
        The values of $H_z^\text{perfect}$, 
            defined along the dashed box surrounding the design area
            (coupling structure), 
            are shown in orange. 
        The values of $H_z^\text{perfect}$ along the top and bottom edges
            of the dashed box are set to zero.
        In this schematic, the fundamental and second-order waveguide modes
            have been chosen as the input and output modes respectively.}
    \label{fig:intro}
\end{figure}
%\section{Objective-first design of waveguide couplers}
% We now demonstrate the objective-first approach
%     by designing waveguide couplers in two-dimensions.
% As such, we do not analyze effects which are only present in full 
%     three-dimensional systems, such as out-of-plane losses.
% Specifically, we work in the two-dimensional transverse electric mode,
%     choosing $H_z$ as the field variable $x$,
%     and $\epsilon^{-1}$ (the inverse of the permittivity) 
%     as the structure variable $p$.
% This results in the following form of the physics residual,
%     \begin{equation}
%     \|g(x,p)\|^2 = 
%     \|g(H_z, \epsilon^{-1})\|^2 = 
%     \| \nabla \times \epsilon^{-1} \nabla \times H_z - \mu_0 \omega^2 H_z \|^2,
%     \end{equation}
%     where $\omega$ is the angular frequency,
%     and $\mu_0$ is the permeability of free-space.
% 
% In order to produce a method applicable 
%     to the design of any linear nanophotonic device,
%     we note that its performance
%     will be completely characterized by the fields along its boundary,
%     and therefore, 
%     we choose a boundary-value formulation for the design objective 
%     as shown in Fig.~\ref{fig:intro}.
% Namely, we first compute the boundary fields needed to obtain
%     perfect performance, $H_z^\text{perfect}$,
%     and then form the design objective as
%     \begin{equation}
%     f(x) = f(H_z) = \begin{bmatrix}
%         H_z - H_z^\text{perfect} \\
%         \frac{\partial H_z}{\partial n} - 
%             \frac{\partial H_z^\text{perfect}}{\partial n}
%         \end{bmatrix}_\text{boundary}
%         = 0,
%     \end{equation}
%     where $\partial H_z / \partial n$ 
%     denotes the spatial derivative in the direction normal to the boundary.
% 
% \begin{figure}[htb]
%     \centering
%     \includegraphics[width=\textwidth]{p3/intro} 
%     \caption{Boundary-value formulation of the design objective.
%         The values of $H_z^\text{perfect}$, 
%             defined along the dashed box surrounding the design area
%             (coupling structure), 
%             are shown in orange. 
%         The values of $H_z^\text{perfect}$ along the top and bottom edges
%             of the dashed box are set to zero.
%         In this schematic, the fundamental and second-order waveguide modes
%             have been chosen as the input and output modes respectively.}
%     \label{fig:intro}
% \end{figure}
%     
Such a design objective is both extremely simple and widely adaptable 
    to the design of linear nanophotonic devices in general.
On the other hand, such a formulation often prohibits the physics residual 
    from disappearing completely; 
    however, we find that very good device performance is still obtained
    as long as the physics residual is sufficiently small.

\section{Implementation}
We employ an alternating directions strategy,
    where both $x$ and $z$ are solved iteratively
    and limit the allowable values of $\epsilon$ to be between
    the permittivity of vacuum and of silicon,
    \begin{equation}
    \epsilon_0 \le \epsilon \le \epsilon_\text{silicon}.
    \end{equation}
Note that a completely binary structure,
    where $\epsilon = \{\epsilon_0, \epsilon_\text{silicon}\}$,
    is needed for manufacturing purposes.
That said, the final designs presented here 
    all have significant portions which are already binary.

\section{Results}
We now demonstrate the objective-first approach
    by designing waveguide couplers in two-dimensions.
As such, we do not analyze effects which are only present in full 
    three-dimensional systems, such as out-of-plane losses.
Specifically, we work in the two-dimensional transverse electric mode,
    choosing $H_z$ as the field variable $x$,
    and $\epsilon^{-1}$ (the inverse of the permittivity) 
    as the structure variable $z$.

Specifically, we present five results which suggest our method
    is indeed able to design arbitrary single-mode, linear nanophotonic devices.
These results comprise of couplers between 
    the following nanophotonic waveguides:
    \begin{enumerate}
    \item coupling between waveguides of different refractive index and width
        (Fig.~\ref{fig:fiber});
    \item coupling between waveguide modes of different order and symmetry
        (Fig.~\ref{fig:mode});
    \item coupling between waveguides that confine light 
        using different principles 
        (index guided vs. distributed Bragg reflection guided), 
        i.e., between a slab waveguide and a photonic crystal fiber
        (Fig.~\ref{fig:aircore})
    \item coupling from a dielectric to a plasmonic metal-insulator-metal 
        waveguide (Fig.~\ref{fig:mim}); and
    \item coupling from a dielectric waveguide to a (plasmonic) metal wire
        (Fig.~\ref{fig:wire}).
    \end{enumerate}
All the designs presented are both highly efficient 
    ($\sim 98\%$ coupling efficiency, in contrast to previous designs
    which achieve only $\sim 70\%$ coupling efficiency\cite{veronis,yang})
    and extremely compact 
    (with footprints of only 1.5-4 square vacuum wavelengths).

Note, however, that the small footprint of the designs presented does not 
    imply small, unresolvable feature sizes.
On the other hand, assuming a vacuum wavelength of 1550 nm,
    the smallest possible feature in any of the examples presented is still
    37 nm $\times$ 37 nm (see Fig.~\ref{fig:fiber}, for example),
    as given by the size of a single grid point;
    placing virtually all designs within reach of current-generation lithographic
    systems.

In addition, good starting points were not required; 
    all initial designs were simply $\epsilon = 9$ everywhere
    (a somewhat arbitrary guess, other values work as well).
Finally, only 15 minutes on a single-core personal computer
    were needed to obtain each design.

In appendix~\ref{ob-1 additional}, we further demonstrate the generality of our method
    by presenting several additional examples including
    \begin{enumerate}
    \item coupling between all four modes of a wide dielectric waveguide;
    \item coupling from a wide, low-index waveguide to the output waveguides in 
        Figs~\ref{fig:mode}-\ref{fig:wire}; and
    \item coupling to selected channels of a set of five plasmonic waveguides.
    \end{enumerate}

Furthermore, appendix~\ref{ob-1 meta} contains designs for metamaterial-like
    devices.

\begin{figure}[htb]
    \centering
    \includegraphics[width=\textwidth]{p3/1} 
    \caption{Coupler from a narrow, high-index ($\epsilon=12.25$) waveguide to
            a wide, low-index ($\epsilon=2.25$) waveguide. 
        The $H_z^\text{perfect}$ boundary values used as 
            the design objective are shown in the upper-left plots, 
            the generated structure is shown in the upper-right plot, and
            the simulated $H_z$ fields of the device are shown 
            in the bottom two plots.
        The computed efficiency of the coupler is high, $99.8\%$, and
            the device is also extremely compact, 
            convering only $36 \times 76$ grid points,
            where the vacuum wavelength is 42 grid points
            (footprint of $1.55$ square vacuum wavelengths).
        Computation time was 15 minutes on a personal computer.}
    \label{fig:fiber}
\end{figure}
\begin{figure}[htb]
    \centering
    \includegraphics[width=\textwidth]{p3/2} 
    \caption{Coupler that converts the fundamental waveguide mode to the
            second-order waveguide mode.
        This problem is quite difficult since the two modes are of 
            opposite symmetry.
        For example, adiabatic approaches cannot be applied to this case.
        However, our method produces a device 
            (which has the same dimensions and vacuum wavelength as 
            Fig.~\ref{fig:fiber})  
            with a coupling efficiency of $98.0\%$. 
        Computation time was 15 minutes on a personal computer.
        }
    \label{fig:mode}
\end{figure}
\begin{figure}[htb]
    \centering
    \includegraphics[width=\textwidth]{p3/3}
    \caption{Coupler between a dielectric slab waveguide to 
            an air-core waveguide.
        Here, not only are the modes of opposite symmetry,
            but the output waveguide operates on a fundamentally different
            principle (distributed reflection) than the input waveguide 
            (index guided).
        The device still achieves an efficiency of $98.9\%$, demonstrating the
            versatility of our method.
        The vacuum wavelength is 25 grid points, 
            while the device footprint is still $36 \times 76$ grid points
            (footprint of 4.38 square vacuum wavelengths).
        Computation time was 15 minutes on a personal computer.
        }
        \label{fig:aircore}
\end{figure}
\begin{figure}[htb]
    \centering
    \includegraphics[width=\textwidth]{p3/4}
    \caption{Coupler between a dielectric slab waveguide to 
            a plasmonic metal-insulator-metal waveguide.
        The efficiency of the device is $97.5\%$ and 
            has the same wavelength and footprint as the device in
            Fig.~\ref{fig:aircore}.
        Computation time was 15 minutes on a personal computer.
        }
        \label{fig:mim}
\end{figure}
\begin{figure}[htb]
    \centering
    \includegraphics[width=\textwidth]{p3/5}
    \caption{Coupler between a dielectric slab waveguide to 
            a plasmonic wire waveguide.
        The efficiency of the device is $99.1\%$ and 
            has the same wavelength and footprint as the device in
            Fig.~\ref{fig:aircore}.
        Computation time was 15 minutes on a personal computer.
        }
        \label{fig:wire}
\end{figure}
% 
% \section{Conclusion}
% We develop a fundamentally new approach to designing physical structures,
%     which we term ``objective-first'', 
%     in that we choose to satisfy the design objective 
%     even above satisfying the physical equation which governs its operation.
% We then apply this approach to the design of 
%     various nanophotonic waveguide couplers,
%     and show that our method produces
%     high-efficiency designs ($\sim 98\%$ efficiency) 
%     in small footprints (1.5-4 square vacuum wavelengths)
%     without needing a good starting point.
% Furthermore, we suggest that such a methodology may be applicable
%     to the design of any single-mode, linear nanophotonic device.\\
% 
% This work has been supported by the 
%     AFOSR MURI for Complex and Robust On-chip Nanophotonics 
%     (Dr. Gernot Pomrenke), grant number FA9550-09-1-0704.
% The Matlab code used is available online\cite{code}.
% \clearpage
% \begin{appendix}