### 2013-04-16 thesis/acknowledgement.tex
Added bibliography and fixed citations.

\prefacesection{Acknowledgement}
Acknowledgement section to be written here.
\afterpreface