### 2013-05-24 thesis/conclusion.tex

\chapter{Conclusion and future work}\label{conclusion}