### 2013-06-04 thesis/acknowledgement.tex
achnowledgement and DABM's corrections.

\prefacesection{Acknowledgements}
\begin{quote}
In the beginning was the Word, and the Word was with God, and the Word was God. He was in the beginning with God. All things were made through him, and without him was not any thing made that was made. In him was life, and the life was the light of men. The light shines in the darkness, and the darkness has not overcome it. (John 1:1-4)
\end{quote}

\begin{quote}This is the message we have heard from him and proclaim to you, that God is light, and in him is no darkness at all. (1 John 1:1) \end{quote}

I have had the privilege, 
    over the course of my studies,
    to become intimately familiar with the study of light 
    both from a \emph{physical} and \emph{spiritual} perspective.
This thesis, not surprisingly, 
    features the equations which describe light.
And when properly considered,
    naturally lead one's mind to the origin of that light, Jesus Christ.

It is the Lord Jesus Christ that I want to acknowledge in this thesis.
I acknowledge Him as my Creator,
    and the One who, in the beginning, said, 
    ```Let there be light', and there was light.'' (Genesis 1:3)
I acknowledge Him as my Savior, the Light of the world,
    who has brought me to God by the forgiveness of my sins in His blood.
He is the ultimate light, 
    apart from which we cannot do or know anything---whether
    we acknowledge Him or not.
And far more significant than any of the progress presented here
    is the work that He promises to do in every willing and submissive heart,
\begin{quote}
For God, who said, ``Let light shine out of darkness,'' has shone in our hearts to give the light of the knowledge of the glory of God in the face of Jesus Christ.
(2 Corinthians 4:6)
\end{quote}

\afterpreface